Succinct Representations for Abstract Interpretation

### 2011-11-05 papers/PLDI2012/implicitization.tex
typos

An additional claim of the methods that distinguish paths inside the loop body \cite{DBLP:conf/pldi/GulwaniZ10,Monniaux_Gonnord_SAS11} is that they tend to generate better invariants that methods that do not, by behaving better with respect to the \emph{widening operators} \cite{CousotCousot_JLC92}
used for enforcing convergence when searching for loop invariants by Kleene iterations. A related technique, \emph{guided static analysis} \cite{DBLP:conf/sas/GopanR07}, computes successive loop invariants for increasing subsets of the transitions taken into account, until all transitions are considered; again, the claim is that this approach avoids some gross over-approximation introduced by widenings.

@@

\subsection{A simple, motivating example}
\label{subsec:rate_lim}
Consider the following program, adapted from \cite{Monniaux_Gonnord_SAS11}, where \lstinline|input($a$, $b$)| stands for a nondeterministic input in $[a,b]$:
\lstinputlisting{rlim_int.c}
This program implements a construct commonly found in control programs (in e.g. automotive or avionics): a rate limiter. For the sake of simplicity, we chose it to be fed a nondeterministic input clamped between $[-100000,100000]$, but in a real system it would be integrated in a reactive control loop and its input connected to a complex system with unknown output range.

@@

The analysis starts by identifying the test taken and not taken during the first iteration of the loop, starting in the loop initialization. The branches not taken are pruned from the loop body, yielding:
\begin{lstlisting}[numbers=none]
  while(1) {
    if(x <= 50) y++;
    else break; /* not taken in phase 1 */
    if(y < 0) break;
    x++;
  }
\end{lstlisting}

@@

Algorithm~ \ref{algo:combined} performs Guided static analysis on the implicitly represented multigraph.

@@

We have implemented our proposed solutions inside a prototype intraprocedural
static analyzer,
as well as the classical abstract interpretation algorithm, and the state-of-the-art
techniques \emph{Path Focusing} \cite{Monniaux_Gonnord_SAS11} and \emph{Guided
Static Analysis} \cite{DBLP:conf/sas/GopanR07}.

>>>>

An additional claim of the methods that distinguish paths inside the loop body \cite{DBLP:conf/pldi/GulwaniZ10,Monniaux_Gonnord_SAS11} is that they tend to generate better invariants than methods that do not, by behaving better with respect to the \emph{widening operators} \cite{CousotCousot_JLC92}
used for enforcing convergence when searching for loop invariants by Kleene iterations. A related technique, \emph{guided static analysis} \cite{DBLP:conf/sas/GopanR07}, computes successive loop invariants for increasing subsets of the transitions taken into account, until all transitions are considered; again, the claim is that this approach avoids some gross over-approximation introduced by widenings.

@@

\subsection{A simple, motivating example}
\label{subsec:rate_lim}
Consider the following program, adapted from \cite{Monniaux_Gonnord_SAS11}, where \lstinline|input($a$, $b$)| stands for a nondeterministic input in $[a,b]$:
\lstinputlisting{rlim_int.c}
This program implements a construct commonly found in control programs (in e.g.
automotive or avionics): a rate limiter. For the sake of simplicity, we chose it
to be fed on a nondeterministic input clamped between $[-100000,100000]$, but in a real system it would be integrated in a reactive control loop and its input connected to a complex system with unknown output range.

@@

The analysis starts by identifying the tests taken and not taken during the first iteration of the loop, starting in the loop initialization. The branches not taken are pruned from the loop body, yielding:
\begin{lstlisting}[numbers=none]
  while(1) {
    if(x <= 50) y++;
    else break; /* not taken in phase 1 */
    if(y < 0) break;
    x++;
  }
\end{lstlisting}

@@

Algorithm~\ref{algo:combined} performs Guided static analysis on the implicitly represented multigraph.

@@

We have implemented our proposed solutions inside a prototype of intraprocedural
static analyzer,
as well as the classical abstract interpretation algorithm, and the state-of-the-art
techniques \emph{Path Focusing} \cite{Monniaux_Gonnord_SAS11} and \emph{Guided
Static Analysis} \cite{DBLP:conf/sas/GopanR07}.
